Data Analysis Extensions

### 2015-07-24 RFCs/HDF5_Library/AnalysisExt/analysis_ext.tex
Add view edits to analysis RFC
Former-commit-id: 98d1a68122164b9cb8ee398b06cbff1087faa882

% Fonts
\usepackage{ifluatex}
\ifluatex
  % Font
  \usepackage{luatextra, xunicode, unicode-math}
  \defaultfontfeatures{Ligatures=TeX}
  \setmainfont[
    UprightFont={*},
    BoldFont={*-Bold},
    ItalicFont={*-Italic},
    BoldItalicFont={*-BoldItalic},
    SlantedFont={*-Italic},
    BoldSlantedFont={*-BoldItalic}
  ]{CMU Serif}
  \setsansfont[
    UprightFont={*},
    BoldFont={*-Bold},
    ItalicFont={*-Oblique},
    BoldItalicFont={*-BoldOblique}
  ]{CMU Sans Serif}
  \setmathfont[math-style=ISO, bold-style=ISO]{Asana Math}
  \setmonofont[
    UprightFont={*-Regular},
    BoldFont={*-Bold},
    ItalicFont={*-Italic},
    BoldItalicFont={*-BoldItalic}
  ]{CMU Typewriter}
\else
  \usepackage{lmodern} % times / lmodern / mathpazo / palatino
  \usepackage[scaled=.95]{helvet}
  \usepackage{courier}
\fi

@@

% Title, author, etc
\title{Data Analysis Extensions}
\author{Jerome Soumagne}
\author{Quincey Koziol}
\date{February 25, 2015}
\rfcversion{2014-07-17.v4}
\revision{July 17, 2014}{Version 1 circulated for comment within The HDF Group.}
\revision{August 20, 2014}{Version 2 circulated for comment within The HDF Group.}
\revision{November 14, 2014}{Version 3 circulated for comment within The HDF Group.}
\revision{February 25, 2015}{Version 4 edits and typo fixing.}

@@

We define in this RFC the components that can enable application developers to
create complex and high-performance queries on both metadata and data elements
within an HDF5 container and retrieve the results of applying those query
operations. Support for these operations on HDF5 containers can be defined via:
\begin{enumerate}
\item New \textit{query} objects\footnote{Query and view objects are
\textit{in-memory} objects, which therefore do not modify the content of the
container.\label{fn:object}} and API routines, enabling the construction of
query requests for execution on HDF5 containers;
\item New \textit{view} objects\footref{fn:object} and API routines, which apply a query to an HDF5
container and return a set of references into the container that fulfills
the query criteria;
\item New \textit{index} objects and API routines, which allows the creation of
indices on the contents of HDF5 containers, to improve query performance.
\end{enumerate}

@@

\section{View Objects}
Applying a query to an HDF5 container creates an HDF5 view object. HDF5 view
objects are run-time, in-memory objects (i.e., not stored in a container) that
consist of read-only references into the contents of the HDF5 container that
the query was applied to. The current API is presented below:
\todo{View API must be revised.}

{\lstsetc
\begin{lstlisting}
hid_t  H5Vcreate(hid_t loc_id, hid_t query_id, hid_t vcpl_id);
herr_t H5Vclose(hid_t view_id);
herr_t H5Vget_location(hid_t view_id, hid_t *loc_id);
herr_t H5Vget_query(hid_t view_id, hid_t *query_id);
herr_t H5Vget_counts(hid_t view_id, hsize_t *attr_count, hsize_t *obj_count,
  hsize_t *elem_region_count);
herr_t H5Vget_attrs(hid_t view_id, hsize_t start, hsize_t count, hid_t attr_id[]);
herr_t H5Vget_objs(hid_t view_id, hsize_t start, hsize_t count, hid_t obj_id[]);
herr_t H5Vget_elem_regions(hid_t view_id, hsize_t start, hsize_t count,
  hid_t dataset_id[], hid_t dataspace_id[]);
\end{lstlisting}
}

View objects are created with \texttt{H5Vcreate}, which applies a query to an
HDF5 container, group hierarchy, or individual object and produces the view
object as a result. The attributes, objects, and/or data elements referenced
within a view can be retrieved by further API calls. Note that currently,
attribute references are not available, this feature will be added in order to
support views that contain attribute references\footnote{See RFC 201X-XX-XX.vX.}.

>>>>

% Fonts
\usepackage{ifluatex}
\ifluatex
  % Font
  \usepackage{luatextra, xunicode, unicode-math}
  \defaultfontfeatures{Ligatures=TeX}
  \setmainfont[
    UprightFont={*},
    BoldFont={*-Bold},
    ItalicFont={*-Italic},
    BoldItalicFont={*-BoldItalic},
    SlantedFont={*-Italic},
    BoldSlantedFont={*-BoldItalic}
  ]{CMU Serif}
  \setsansfont[
    UprightFont={*},
    BoldFont={*-Bold},
    ItalicFont={*-Oblique},
    BoldItalicFont={*-BoldOblique}
  ]{CMU Sans Serif}
  \setmathfont[math-style=ISO, bold-style=ISO]{Asana Math}
  \setmonofont[
    UprightFont={*-Regular},
    BoldFont={*-Bold},
    ItalicFont={*-Italic},
    BoldItalicFont={*-BoldItalic}
  ]{CMU Typewriter}
\else
  \usepackage{times} % times / lmodern / mathpazo / palatino
  \usepackage[scaled=.95]{helvet}
  \usepackage{courier}
\fi

@@

\usepackage[margins]{trackchanges}
\addeditor{JS}
\addeditor{QK}
% annote, note, add, remove, change

% Title, author, etc
\title{Data Analysis Extensions}
\author{Jerome Soumagne}
\author{Quincey Koziol}
\date{July 24, 2015}
\rfcversion{2014-07-17.v4}
\revision{Jul. 17, 2014}{Version 1 circulated for comment within The HDF Group.}
\revision{Aug. 20, 2014}{Version 2 circulated for comment within The HDF Group.}
\revision{Nov. 14, 2014}{Version 3 circulated for comment within The HDF Group.}
\revision{Feb. 25, 2015}{Version 4 edits and typo fixing.}
\revision{Jul. 24, 2015}{Version 5 with edits to view object.}

@@

We define in this RFC the components that can enable application developers to
create complex and high-performance queries on both metadata and data elements
within an HDF5 container and retrieve the results of applying those query
operations. Support for these operations on HDF5 containers can be defined via:
\begin{enumerate}
\item New \textit{query} objects\footnote{Query and view objects are
\textit{in-memory} objects, which therefore do not modify the content of the
container.\label{fn:object}} and API routines, enabling the construction of
query requests for execution on HDF5 containers;
\item New \textit{view} objects\footref{fn:object} and API routines, which
\change[JS]{apply a query to an HDF5 container and return }{are the result of a query applied
to an HDF5 container and consist of} a set of references into the
container that fulfills the query criteria;
\item New \textit{index} objects and API routines, which allows the creation of
indices on the contents of HDF5 containers, to improve query performance.
\end{enumerate}

@@

\section{View Objects}
Applying a query to an HDF5 container creates an HDF5 view. HDF5 view
objects are run-time, in-memory objects (i.e., not stored in a container)
\add[JS]{but which can be persisted}, that
consist of read-only references \add[JS]{to attributes, objects and data regions}
into the contents of the HDF5 container that the query was applied to.
\add[JS]{The view object can be therefore seen as a virtual container. It is
defined as an anonymous HDF5 group, which consists of three predefined datasets that
contain attribute, object and region references. A new view is created by the
following routine, which applies a query to an HDF5 container, group hierarchy,
or individual object and returns the object ID of the newly created group:}
{\lstsetc
\begin{lstlisting}
hid_t H5Qapply(hid_t loc_id, hid_t query_id, hid_t vcpl_id, unsigned *result);
\end{lstlisting}
}

\add[JS]{As mentioned, the created view may also need to be persisted. For
coherency, a time stamp may be attached to it so that its states has a meaning
compared to the state of the container that the query was applied to (as the
container may have been modified in the meantime). It may also be useful in
that case to define different states to the view (\textit{dead} or
\textit{live}) so that the user knows whether the view is current or not.}

To further ease the retrieval of the results from the view and for convenience,
an API specific to the view object may be defined:

\note[JS]{Which ones do we really need???}
{\lstsetc
\begin{lstlisting}
herr_t H5Vget_location(hid_t view_id, hid_t *loc_id);
herr_t H5Vget_query(hid_t view_id, hid_t *query_id);
herr_t H5Vget_counts(hid_t view_id, hsize_t *attr_count, hsize_t *obj_count,
  hsize_t *elem_region_count);
herr_t H5Vget_attrs(hid_t view_id, hsize_t start, hsize_t count, hid_t attr_id[]);
herr_t H5Vget_objs(hid_t view_id, hsize_t start, hsize_t count, hid_t obj_id[]);
herr_t H5Vget_elem_regions(hid_t view_id, hsize_t start, hsize_t count,
  hid_t dataset_id[], hid_t dataspace_id[]);
\end{lstlisting}
}

\remove[JS]{View objects are created with \texttt{H5Vcreate}, which applies a query to an
HDF5 container, group hierarchy, or individual object and produces the view
object as a result. The attributes, objects, and/or data elements referenced
within a view can be retrieved by further API calls.} Note that currently,
attribute references are not available, this feature will be added in order to
support views that contain attribute references\footnote{See RFC 201X-XX-XX.vX.}.
